---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Volatility Spillovers from US to SA Markets"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Ruan Geldenhuys" # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "22550801\\@sun.ac.za" # First Author's Email address

Author2: "Nico Katzke"
Ref2: "Stellenbosch University, Stellenbosch, South Africa"
Email2: "nfkatzke\\@gmail.com"
CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

# If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
keywords: "Multivariate GARCH \\sep Spillovers" # Use \\sep to separate

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  I investigate the relationship between the voaltilities of S\\&P 500 and the JSE Top 40. The purpose of this study is to investigate if this relationship changes in any significant way during the two biggest crisis periods in the last two decades, namely the Global Financial Crisis and Covid-19. I first do a stratification analysis which reveals significant evidence of these two indices sharing periods of high volatility. I then fit multiple multivariate GARCH models to further investigate the volatility relationship and find... 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(tidyverse)
library(readxl)
library(fmxdat)
library(tseries)
library(knitr)
library(MTS)
library(zoo)
library(ggthemes)
library(rmgarch)
library(rugarch)
library(mgarchBEKK)
library(tikzDevice)
library(moments)
library(xtable)

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

```

<!-- ############################## -->

<!-- # Start Writing here: -->

<!-- ############################## -->

# Introduction \label{Introduction}


# Data and Methodology

## Data

Three return series are used in the analysis that follows. These are the monthly returns for the S\\&P 500 and the JSE Top 40, as well as thhe ZAR/USD exchange rate. The exchange rate is represented as the amount of Rands required to buy one US Dollar. Since the series is represented as a growth rate, a postive growth rate represents a depreciation of the Rand, and conversely, an appreciation of the Dollar. The returns for these 3 series' are visualised below in Figures \ref{Figure1} to \ref{Figure3}. 

```{r Figure1, warning =  FALSE, fig.align = 'center', fig.cap = "S\\&P 500 Returns \\label{Figure1}", fig.ext = 'png', fig.height = 5, fig.width = 6}

global_indices <- readRDS("data/Global_Indices.rds")
local_indices <- readRDS("data/LCL_Indices.rds")
USDZAR <- readRDS("data/USDZAR.rds")

SP <- global_indices %>% #This includes rand returns
    filter(Tickers == "SPXT") %>% 
    select(c(date, Returns)) %>% 
    rename(SP500 = Returns)

lcl_index <- "J200" # I create this variable so the choice of SA index can easily be changed
JSE <- local_indices %>% 
    filter(Tickers == lcl_index) %>% 
    select(c(date, Returns)) %>% 
    rename(JSE40 = Returns)

joinedDF <- left_join(SP, JSE, by = 'date')

firstdate <- joinedDF %>%  slice(1) %>% pull(date)

ZARUSD <- USDZAR %>% 
    select(c(date, value)) %>% 
    filter(date >= firstdate) %>% 
    mutate(yearmonth = format(ymd(date), "%Y-%m")) %>% 
    group_by(yearmonth) %>% 
    mutate(ZARUSD = dplyr::last(value)/dplyr::first(value) - 1) %>% 
    filter(date == dplyr::last(date)) %>% 
    ungroup() %>% 
    slice(-1) %>% 
    select(c(date, ZARUSD))

joinedDF <- left_join(ZARUSD, joinedDF, by = 'date') 
joinedDF <- joinedDF[c(1,3,4,2)]

#Plot the returns
returnplots <- returns_plotter(joinedDF, c("S&P 500", "JSE Top 40", "USD/ZAR"))
returnplots$`S&P 500`


```

```{r Figure2, warning =  FALSE, fig.align = 'center', fig.cap = "JSE Top 40 Returns \\label{Figure2}", fig.ext = 'png', fig.height = 5, fig.width = 6}

returnplots$`JSE Top 40`

```

Not much information can be revealed through simply observing the returns over time. However, when investigating the squared returns as a measure of volatility, it is clear to see that the JSE Top 40 is substantially more volatile than the S\\&P 500. This result is reinforced by Table \ref{tab1}, where the JSE showcases a standard deviation considerably higher than that of the S\\&P. Interestingly, the JSE Top 40 showcases higher average monthly returns, however that comes at the cost of the increased volatility as described above. Lastly, as shown in Table \ref{tab1} the S\\&P experienced the largest draw down, while the JSE experienced the largest uptick.

```{r Figure3, warning =  FALSE, fig.align = 'center', fig.cap = "ZAR/USD Returns \\label{Figure3}", fig.ext = 'png', fig.height = 5, fig.width = 6}

returnplots$`USD/ZAR`

```


```{r, results='asis'}
sumstats <- function(vec){
    mean <- mean(vec)
    median <- median(vec)
    sd <- sd(vec)
    kurtosis <- kurtosis(vec)
    skewness <- skewness(vec)
    minimum <- min(vec)
    maximum <- max(vec)
    
    result_vec <- c(mean, median, sd, kurtosis, skewness, minimum, maximum)
    return(result_vec)
}

SP_stats <- sumstats(joinedDF$SP500)
JSE_stats <- sumstats(joinedDF$JSE40)
ZARUSD_stats <- sumstats(joinedDF$ZARUSD)

stats_df <- data.frame(
    Measure = c("Mean", "Median", "Std. Dev.", "Kurtosis", "Skewness", "Minimum", "Maximum"),
    SP500 = SP_stats,
    JSE40 = JSE_stats,
    ZARUSD = ZARUSD_stats
)
colnames(stats_df) <- c("", "S&P 500", "JSE Top 40", "ZAR/USD")

table <- xtable(stats_df, caption = "Summary Statistics \\label{tab1}", digits = c(0,0,4,4,4))

  print.xtable(table,
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H',
             # scalebox = 0.3,
             comment = FALSE,
             caption.placement = 'top',
             include.rownames = FALSE
             )

```


## Methodology 

# Results

## DCC

## Go-GARCH

## BEKK-GARCH


# Conclusion

I hope you find this template useful. Remember, stackoverflow is your friend - use it to find answers to questions. Feel free to write me a mail if you have any questions regarding the use of this package. To cite this package, simply type citation("Texevier") in Rstudio to get the citation for @Texevier (Note that uncited references in your bibtex file will not be included in References).

<!-- Make title of bibliography here: -->

<!-- \newpage -->

\newpage

# References {.unnumbered}

::: {#refs}
:::

# Appendix {.unnumbered}

## Appendix A {.unnumbered}

Some appendix information here

## Appendix B {.unnumbered}
